A Comparison of R Tools for Nonlinear Least Squares Modeling

### 2023-11-13 JNABresubmission20231101/JNAB2review-response.txt

On page 4 the authors state that if a selfStart model is not available nls() sets the starting
values to 1. That may be the case now but the original intention was certainly that a user
calling nls() would provide starting estimates suitable to the model and data.

Also on page 4 I don’t understand the comment in the hobbs.res function that This variant
uses looping. I don’t see any explicit loops. There will naturally be looping in the C code
that implements the arithmetic operations on vectors but that is inevitable.

>> The comment was a left-over from an earlier variant of the functions and should have been
>> removed. Done. Thanks for spotting this.
>> We also noted that the model is UNscaled, and edited the comment.
>> An explanatory comment has been added.
>> 
>> The particular form of these explicit residual and Jacobian functions comes from
>> their translation from BASIC codes of the 1970s, as adapted in @jnmws87. The use
>> of the Laplace form of the residual and the inclusion of data within the functions
>> reflects choices of that era that are at odds with current practice. Some users
>> do still want or need to provide problems as explicit functions, particularly
>> for problems that are NOT regressions. For example, the Rosenbrock banana-valley
>> test problem can be provided this way, where the two "residuals" are different
>> functional forms.

>> In the example output, we have added a printout of the calling sequence to document 
>> the results better.

@@

>> In October 2023, one of us (JN) had a query on usage of this formulation where 
>> performance was important. Moreover, the query asked how to extract standard errors
>> and has led to a minor but important addition to the "Intro to nlsr" vignette.
>> Generally we agree that users today are much less likely to use explicit code.

Also, this is the first time that the authors have invoked their non-standard definition of
“residual” as the negative of what most others mean by “residual”, and they haven’t mentioned
this. I would discourage the authors from using an idiosyncratic definition of “residual” simply
to save writing a negative sign in a formula for the Jacobian.

>> We believe our comment above addresses this. We have explained our reasoning for the use
>> of the Laplace residual in other works, but present usage or commentary is a red herring
>> that is irrelevant to today's software and practice except for a note in the hobbs.res
>> function definition. Indeed, the main advance that package 'nlsr' represents is to 
>> automatically provide analytic derivatives that caused so much work and introduced 
>> so many errors in the 1970s and 1980s.

>> We have also removed the sentences (just after the definition of the Logistic models):
>> "The residuals can be written either as r1=y−model(p) or r=model(p)−y since their sum of 
>> squares has an identical value. The second form avoids a potential sign error if we need 
>> to evaluate derivatives."
>>    but added the definition of the residual as r = y - model(p).

@@

>> Other notes:
>>
>> In re-reading our paper, made some minor wording changes to avoid what we believe may be
>> confusing. In particular, we reworded the explanation of the `y` vector for gsl_nls() 
>> under "Functional specification of problems" to accord better with the man page.

>> We noticed the citation to the "illustrative example" lacked a "p." to indicate the
>> page number. We added "csl: ./apa.csl" to the YAML header and downloaded it and have
>> included this file. The output of rendering the Rmd file asks that custom CSL not be
>> used. We are not sure how to resolve this?

>> We have also changed all "NOT" occurrences to simply "not", and removed some boldface
>> where it is unnecessary.

>> A minor clarification of the iteration limit termination tests has been added, namely,
>>  "examination of the code for the different algorithms" since the different `nls()`
>> algorithms may use different tests.

### 2023-11-13 JNABresubmission20231101/ComparingNLStools.Rmd
resub 231101

draft: true
author:
    - name: John C. Nash
      affiliation: retired professor, University of Ottawa
      address:
      - Telfer School of Management
      - Ottawa ON Canada K1N 6N5
      orcid: 0000-0002-2762-8039
      email: profjcnash@gmail.com
    - name: Arkajyoti Bhattacharjee
      affiliation: Indian Institute of Technology
      address:
      - Department of Mathematics and Statistics
      - Kanpur
      email: arkastat98@gmail.com
type: package
output: 
  rjtools::rjournal_web_article:
      self_contained: yes
      toc: no
  # pdf_document
bibliography: ./ComparingNLStools.bib
csl: ./apa.csl
---

@@

`optimx` (@p-optimx) -- This provides 
optimizers that can be applied to minimize a nonlinear function which could be
a nonlinear sum of 
squares. not generally recommended if nonlinear 
least squares programs can be easily used, but provides a check and alternative
solvers.

@@

$$   Asym =  b_1 = 100 * c_1 $$
$$ exp(xmid/scal)  =  b_2 = 10 * c_2 $$
$$ 1/scal  =  b_3 = 0.1*c3 $$
To allow for simpler discussion, let us say that the parameters form a (named) vector
$p$ and the model function is called $model(p)$. The residuals can be written
either as $r_1 = y - model(p)$ or $r = model(p) - y$
since their sum of squares has an identical value. The second form
avoids a potential sign error if we need to evaluate derivatives.

@@

```{r ex02nlsr}
unlx1<-try(nlxb(formula=frmu, start=stu1, data=weeddf))
cat("unlx1:")
print(unlx1) 
snlx1<-try(nlxb(formula=frms, start=sts1, data=weeddf))
pshort(snlx1) # a short-form output
tnlx1<-try(nlxb(formula=frmt, start=stt1, data=weeddf))
pshort(tnlx1) # alternatively print(tnlx1)
```

@@

```{r ex02minpack}
unlm1<-try(nlsLM(formula=frmu, start=stu1, data=weeddf))
pnls(unlm1)  # Short form of output
snlm1<-try(nlsLM(formula=frms, start=sts1, data=weeddf))
pnls(snlm1)
tnlm1<-try(nlsLM(formula=frmt, start=stt1, data=weeddf))
pnls(tnlm1) # short form to give sum of squares, else use summary(tnlm1)
```

@@

```{r ex02gslnls}
library(gslnls)
ugslnls1<-try(gsl_nls(fn = frmu, data = weeddf,  start = stu1))
pnls(ugslnls1) # to get sum of squares
sgslnls1<-try(gsl_nls(fn = frms, data = weeddf,  start = sts1))
pnls(sgslnls1) # Use summary() to get display
tgslnls1<-try(gsl_nls(fn = frmt, data = weeddf,  start = stt1))
pnls(tgslnls1) 
```

@@

```{r ex05, echo=TRUE}  
stspecial <- c(Asym = 35.532,  xmid = 43376,  scal = -2935.4)
badstart <- nlxb(formula = frmt, start = stspecial, data = weeddf)
print(badstart)
```

@@

```{r exhobbsfn, eval=TRUE}
hobnlfb<-nlfb(start=stu1, resfn=hobbs.res, jacfn=hobbs.jac)
pshort(hobnlfb) # use print(hobnlfb) for more detail
hobnlm<-nls.lm(par=stu1, fn=hobbs.res, jac=hobbs.jac)
pnlslm(hobnlm)  
hobgsln<-gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12))
pnls(hobgsln)
hobgsl<-gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12), jac=hobbs.jac)
pnls(hobgsl) # using analytic Jacobian
```

@@

Termination after what may be considered excessive computation is also important.
`nls()` stops after `maxiter` "iterations". The meaning of "iteration" may require an 
examination of the code. We terminate execution when the number of residual or 
Jacobian evaluations exceed set limits. Generally, we prefer larger limits than the 
default `maxiter = 50` of `nls()` to avoid stopping early.

@@

```{r ex10, echo=TRUE} 
# Start MUST be feasible i.e. on or within bounds
anlshob1b <- nls(frms, start = sts1, data = weeddf, lower = c(0,0,0),
             upper = c(2,6,3), algorithm = 'port')
pnls(anlshob1b) #  check the answer (short form)
# nlsLM seems NOT to work with bounds in this example
anlsLM1b <- nlsLM(frms, start = sts1, data = weeddf, lower = c(0,0,0), upper = c(2,6,3))
pnls(anlsLM1b)
# also no warning if starting out of bounds, but gets a good answer!!
st4 <- c(c1 = 4, c2 = 4, c3 = 4)
anlsLMob <- nlsLM(frms, start = st4, data = weeddf, lower = c(0,0,0), upper = c(2,6,3))
pnls(anlsLMob)
# Try nlsr::nlxb()
anlx1b <- nlxb(frms, start = sts1, data = weeddf, lower = c(0,0,0), upper = c(2,6,3))
pshort(anlx1b)
```

@@

```{r ex10m, echo=TRUE}
# Hobbsmaskx.R -- masks with formula specification of the problem
require(nlsr); require(minpack.lm); traceval <- FALSE
stu <- c(b1 = 200, b2 = 50, b3 = 0.3) # a default starting vector (named!)
sts <- c(c1 = 2, c2 = 5, c3 = 3) # a default scaled starting vector (named!)
# fix first parameter
anxbmsk1 <- try(nlxb(frmu, start = stu, data = weeddf, lower = c(200, 0, 0), 
			upper = c(200, 60, 3), trace=traceval))
print(anxbmsk1)
anlM1 <- try(nlsLM(frmu, start = stu, data = weeddf, lower = c(200, 0, 0), 
			upper=c(200, 60, 3), trace = traceval))
pnls(anlM1)
anlsmsk1 <- try(nls(frmu, start = stu, data = weeddf, lower = c(200, 0, 0), 
		upper = c(200, 60, 3),  algorithm = "port", trace = traceval))
pnls(anlsmsk1)
# Hobbs scaled problem with bounds, formula specification
anlxmsks1 <- nlxb(frms, start = sts, data = weeddf, lower = c(2, 0, 0),
                  upper = c(2, 6, 30))
print(anlxmsks1)
anlshmsk1 <- nls(frms, start = sts, trace = traceval, data = weeddf, 
                 lower = c(2, 0, 0), upper = c(2, 6, 30), algorithm = 'port')
pnls(anlshmsk1)
anlsLMmsks1 <- nlsLM(frms, start = sts, data = weeddf, lower = c(2,0,0),
                 upper = c(2,6,30))
pnls(anlsLMmsks1)

# Test with all parameters masked
anlxmskall <- try(nlxb(frms, start=sts, data=weeddf, lower=sts, upper=sts))
print(anlxmskall)
```

@@

`nls()` can be called without specifying the `data` argument. In this case, it will
search in the available environments (i.e., workspaces) for suitable data objects. 
We do not like this approach, but it is "the R way". R allows users to leave many 
objects in the default (`.GlobalEnv`) workspace. Moreover, users have to actively 
suppress saving this workspace (`.RData`) on exit; otherwise any such file in 
the path will be loaded on startup. R users in our
acquaintance avoid saving the workspace because of lurking data and
functions that may cause unwanted results.

@@

```{r nlswtx, echo=TRUE}
wts <- 0.5 ^ tt # simple weights
frmlogis <- weed ~ Asym / (1 + exp((xmid - tt)/scal))
Asym <- 1; xmid <- 1; scal <- 1
nowt <- nls(weed ~ SSlogis(tt, Asym, xmid, scal)) # UNWEIGHTED
rnowt <- nowt$m$resid() # This has UNWEIGHTED residual and Jacobian. Does NOT take coefficients.
attr(rnowt, "gradient") <- NULL; rnowt
usewt <- nls(weed ~ SSlogis(tt, Asym, xmid, scal), weights = wts)
rusewt <- usewt$m$resid() # WEIGHTED. Does NOT take coefficients.
attr(rusewt, "gradient") <- NULL; rusewt
## source("nlsModel.R") # or use {r nlsmodelsource, echo=FALSE} code chunk
nmod0 <- nlsModel(frmlogis, data = weeddf, start = c(Asym = 1, xmid = 1, scal = 1), wts = wts)
rn0 <- nmod0$resid() # Parameters are supplied in nlsModel() `start` above.
attr(rn0, "gradient") <- NULL; rn0 # weighted residuals at starting coefficients
nmod <- nlsModel(frmlogis, data = weeddf, start = coef(usewt), wts = wts)
rn <- nmod$resid()
attr(rn,"gradient")<-NULL; rn # same as rusewt
```

@@

```{r tetrarun}
time <- c( 1,  2,  3,  4,  6 , 8, 10, 12, 16)
conc <- c( 0.7, 1.2, 1.4, 1.4, 1.1, 0.8, 0.6, 0.5, 0.3)
NLSdata <- data.frame(time,conc)
NLSstart <- c(lrc1 = -2, lrc2 = 0.25, A1 = 150, A2 = 50) # a starting vector (named!)
NLSformula <- conc ~ A1 * exp(-exp(lrc1) * time) + A2 * exp(-exp(lrc2) * time)
tryit <- try(nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE))
print(tryit)
```

@@

```{r log4ways, echo=TRUE, eval=FALSE}
DNase1 <- subset(DNase, Run == 1) # select the data
## using a selfStart model - do not specify the starting parameters
fm1 <- nls(density ~ SSlogis(log(conc), Asym, xmid, scal), DNase1)
summary(fm1)

## using conditional linearity - leave out the Asym parameter
fm2 <- nls(density ~ 1 / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(xmid = 0, scal = 1),
                 algorithm = "plinear")
summary(fm2)

## without conditional linearity
fm3 <- nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1,
                 start = list(Asym = 3, xmid = 0, scal = 1))
summary(fm3)

## using Port's nl2sol algorithm
fm4 <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(Asym = 3, xmid = 0, scal = 1),
                 algorithm = "port"))
summary(fm4)

## using conditional linearity AND Asym does not work
fm2a <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)), 
                 data = DNase1, start = list(Asym=3, xmid = 0, scal = 1),
                 algorithm = "plinear", trace = TRUE))
summary(fm2a)
```

>>>>

draft: true
author:
    - name: John C. Nash
      affiliation: retired professor, University of Ottawa
      address:
      - Telfer School of Management
      - Ottawa ON Canada K1N 6N5
      orcid: 0000-0002-2762-8039
      email: profjcnash@gmail.com
    - name: Arkajyoti Bhattacharjee
      affiliation: Indian Institute of Technology
      address:
      - Department of Mathematics and Statistics
      - Kanpur
      email: arkastat98@gmail.com
type: package
output: 
  rjtools::rjournal_article:
    self_contained: yes
    toc: no
bibliography: ./ComparingNLStools.bib
---

@@

`optimx` (@p-optimx) -- This provides 
optimizers that can be applied to minimize a nonlinear function which could be
a nonlinear sum of squares. Not generally recommended if nonlinear 
least squares programs can be easily used, but provides a check and alternative
solvers.

@@

$$   Asym =  b_1 = 100 * c_1 $$
$$ exp(xmid/scal)  =  b_2 = 10 * c_2 $$
$$ 1/scal  =  b_3 = 0.1*c3 $$
To allow for simpler discussion, let us say that the parameters form a (named) vector
$p$ and the model function is called $model(p)$ with $r = y - model(p)$.

@@

```{r ex02nlsr}
unlx1<-try(nlxb(formula=frmu, start=stu1, data=weeddf))
if (! inherits(unlx1, "try-error")) {cat("unlx1:"); print(unlx1) }
snlx1<-try(nlxb(formula=frms, start=sts1, data=weeddf))
if (! inherits(snlx1, "try-error"))pshort(snlx1) # a short-form output
tnlx1<-try(nlxb(formula=frmt, start=stt1, data=weeddf))
if (! inherits(tnlx1, "try-error"))pshort(tnlx1) # alternatively print(tnlx1)
```

@@

```{r ex02minpack}
unlm1<-try(nlsLM(formula=frmu, start=stu1, data=weeddf))
if (! inherits(unlm1, "try-error"))pnls(unlm1)  # Short form of output
snlm1<-try(nlsLM(formula=frms, start=sts1, data=weeddf))
if (! inherits(snlm1, "try-error"))pnls(snlm1)
tnlm1<-try(nlsLM(formula=frmt, start=stt1, data=weeddf))
if (! inherits(tnlm1, "try-error"))pnls(tnlm1) # short form to give sum of squares, else use summary(tnlm1)
```

@@

```{r ex02gslnls}
library(gslnls)
ugslnls1<-try(gsl_nls(fn = frmu, data = weeddf,  start = stu1))
if (! inherits(ugslnls1, "try-error")) pnls(ugslnls1) # to get sum of squares
sgslnls1<-try(gsl_nls(fn = frms, data = weeddf,  start = sts1))
if (! inherits(sgslnls1, "try-error")) pnls(sgslnls1) # Use summary() to get display
tgslnls1<-try(gsl_nls(fn = frmt, data = weeddf,  start = stt1))
if (! inherits(tgslnls1, "try-error")) pnls(tgslnls1) 
```

@@

```{r ex05, echo=TRUE}  
stspecial <- c(Asym = 35.532,  xmid = 43376,  scal = -2935.4)
badstart <- try(nlxb(formula = frmt, start = stspecial, data = weeddf))
if (! inherits(badstart, "try-error")) print(badstart)
```

@@

```{r exhobbsfn, eval=TRUE}
cat("hobnlfb<-nlfb(start=stu1, resfn=hobbs.res, jacfn=hobbs.jac)\n")
hobnlfb<-try(nlfb(start=stu1, resfn=hobbs.res, jacfn=hobbs.jac))
if (! inherits(hobnlfb, "try-error")) pshort(hobnlfb) # use print(hobnlfb) for more detail
cat("hobnlm<-nls.lm(par=stu1, fn=hobbs.res, jac=hobbs.jac)\n")
hobnlm<-try(nls.lm(par=stu1, fn=hobbs.res, jac=hobbs.jac))
if (! inherits(hobnlm, "try-error")) pnlslm(hobnlm)
cat("hobgsln<-gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12))\n")
hobgsln<-try(gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12)))
if (! inherits(hobgsln, "try-error")) pnls(hobgsln)
cat("hobgsl<-gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12), jac=hobbs.jac)\n")
hobgsl<-try(gsl_nls(start=stu1, fn=hobbs.res, y=rep(0,12), jac=hobbs.jac))
if (! inherits(hobgsl, "try-error")) pnls(hobgsl) # using analytic Jacobian
```

@@

Termination after what may be considered excessive computation is also important.
`nls()` stops after `maxiter` "iterations". The meaning of "iteration" may require an 
examination of the code for the different algorithms. `nlsr` terminates execution 
when the number of residual or Jacobian evaluations exceed set limits. Generally, 
we prefer larger limits than the default `maxiter = 50` of `nls()` to avoid stopping 
early, though this may result in some unnecessary computations.

@@

```{r ex10, echo=TRUE} 
# Start MUST be feasible i.e. on or within bounds
anlshob1b <- try(nls(frms, start = sts1, data = weeddf, lower = c(0,0,0),
             upper = c(2,6,3), algorithm = 'port'))
if (! inherits(anlshob1b, "try-error")) pnls(anlshob1b) #  check the answer (short form)
# nlsLM seems NOT to work with bounds in this example
anlsLM1b <- try(nlsLM(frms, start = sts1, data = weeddf, lower = c(0,0,0), upper = c(2,6,3)))
if (! inherits(anlsLM1b, "try-error")) pnls(anlsLM1b)
# also no warning if starting out of bounds, but gets a good answer!!
st4 <- c(c1 = 4, c2 = 4, c3 = 4)
anlsLMob <- try(nlsLM(frms, start = st4, data = weeddf, lower = c(0,0,0), upper = c(2,6,3)))
if (! inherits(anlsLMob, "try-error")) pnls(anlsLMob)
# Try nlsr::nlxb()
anlx1b <- try(nlxb(frms, start = sts1, data = weeddf, lower = c(0,0,0), upper = c(2,6,3)))
if (! inherits(anlx1b, "try-error")) pshort(anlx1b)
```

@@

```{r ex10m, echo=TRUE}
# Hobbsmaskx.R -- masks with formula specification of the problem
require(nlsr); require(minpack.lm); traceval <- FALSE
stu <- c(b1 = 200, b2 = 50, b3 = 0.3) # a default starting vector (named!)
sts <- c(c1 = 2, c2 = 5, c3 = 3) # a default scaled starting vector (named!)
# fix first parameter
anxbmsk1 <- try(nlxb(frmu, start = stu, data = weeddf, lower = c(200, 0, 0), 
			upper = c(200, 60, 3), trace=traceval))
if (! inherits(anxbmsk1, "try-error")) print(anxbmsk1)
anlM1 <- try(nlsLM(frmu, start = stu, data = weeddf, lower = c(200, 0, 0), 
			upper=c(200, 60, 3), trace = traceval))
if (! inherits(anlM1, "try-error")) pnls(anlM1)
anlsmsk1 <- try(nls(frmu, start = stu, data = weeddf, lower = c(200, 0, 0), 
		upper = c(200, 60, 3),  algorithm = "port", trace = traceval))
if (! inherits(anlsmsk1, "try-error")) pnls(anlsmsk1)
# Hobbs scaled problem with bounds, formula specification
anlxmsks1 <- try(nlxb(frms, start = sts, data = weeddf, lower = c(2, 0, 0),
                  upper = c(2, 6, 30)))
if (! inherits(anlxmsks1, "try-error")) print(anlxmsks1)
anlshmsk1 <- try(nls(frms, start = sts, trace = traceval, data = weeddf, 
                 lower = c(2, 0, 0), upper = c(2, 6, 30), algorithm = 'port'))
if (! inherits(anlshmsk1, "try-error")) pnls(anlshmsk1)
anlsLMmsks1 <- try(nlsLM(frms, start = sts, data = weeddf, lower = c(2,0,0),
                 upper = c(2,6,30)))
if (! inherits(anlsLMmsks1, "try-error")) pnls(anlsLMmsks1)

# Test with all parameters masked
anlxmskall <- try(nlxb(frms, start=sts, data=weeddf, lower=sts, upper=sts))
if (! inherits(anlxmskall, "try-error")) print(anlxmskall)
```

@@

`nls()` can be called without specifying the `data` argument. In this case, it will
search in the available environments (i.e., workspaces) for suitable data objects. 
We do not like this approach, but it is "the R way". R allows users to leave many 
objects in the default (`.GlobalEnv`) workspace. Moreover, users have to actively 
suppress saving this workspace (`.RData`) on exit, otherwise any such file in 
the path will be loaded on startup. R users in our
acquaintance avoid saving the workspace because of lurking data and
functions that may cause unwanted results.

@@

```{r nlswtx, echo=TRUE}
wts <- 0.5 ^ tt # simple weights
frmlogis <- weed ~ Asym / (1 + exp((xmid - tt)/scal))
Asym <- 1; xmid <- 1; scal <- 1
nowt <- try(nls(weed ~ SSlogis(tt, Asym, xmid, scal))) # UNWEIGHTED
if (! inherits(nowt, "try-error")) {
  rnowt <- nowt$m$resid() # This has UNWEIGHTED residual and Jacobian. Does NOT take coefficients.
  attr(rnowt, "gradient") <- NULL 
} else rnowt <- NULL
rnowt
usewt <- try(nls(weed ~ SSlogis(tt, Asym, xmid, scal), weights = wts))
if (! inherits(usewt, "try-error")) {
  rusewt <- usewt$m$resid() # WEIGHTED. Does NOT take coefficients.
  attr(rusewt, "gradient") <- NULL
} else rusewt <- NULL
rusewt
## source("nlsModel.R") # or use {r nlsmodelsource, echo=FALSE} code chunk
nmod0 <- nlsModel(frmlogis, data = weeddf, start = c(Asym = 1, xmid = 1, scal = 1), wts = wts)
rn0 <- nmod0$resid() # Parameters are supplied in nlsModel() `start` above.
attr(rn0, "gradient") <- NULL; rn0 # weighted residuals at starting coefficients
nmod <- nlsModel(frmlogis, data = weeddf, start = coef(usewt), wts = wts)
rn <- nmod$resid()
attr(rn,"gradient")<-NULL; rn # same as rusewt
```

@@

```{r tetrarun, echo=TRUE}
time <- c( 1,  2,  3,  4,  6 , 8, 10, 12, 16)
conc <- c( 0.7, 1.2, 1.4, 1.4, 1.1, 0.8, 0.6, 0.5, 0.3)
NLSdata <- data.frame(time,conc)
NLSstart <- c(lrc1 = -2, lrc2 = 0.25, A1 = 150, A2 = 50) # a starting vector (named!)
NLSformula <- conc ~ A1 * exp(-exp(lrc1) * time) + A2 * exp(-exp(lrc2) * time)
tryit <- try(nls(NLSformula, data = NLSdata, start = NLSstart, trace = TRUE))
if (! inherits(tryit, "try-error")) print(tryit)
```

@@

```{r log4ways, echo=TRUE, eval=FALSE}
DNase1 <- subset(DNase, Run == 1) # select the data
## using a selfStart model - do not specify the starting parameters
fm1 <- try(nls(density ~ SSlogis(log(conc), Asym, xmid, scal), DNase1))
if (! inherits(fm1, "try-error")) summary(fm1)

## using conditional linearity - leave out the Asym parameter
fm2 <- try(nls(density ~ 1 / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(xmid = 0, scal = 1),
                 algorithm = "plinear"))
if (! inherits(fm2, "try-error")) summary(fm2)

## without conditional linearity
fm3 <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1,
                 start = list(Asym = 3, xmid = 0, scal = 1)))
if (! inherits(fm3, "try-error")) summary(fm3)

## using Port's nl2sol algorithm
fm4 <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)),
                 data = DNase1, start = list(Asym = 3, xmid = 0, scal = 1),
                 algorithm = "port"))
if (! inherits(fm4, "try-error")) summary(fm4)

## using conditional linearity AND Asym does not work
fm2a <- try(nls(density ~ Asym / (1 + exp((xmid - log(conc)) / scal)), 
                 data = DNase1, start = list(Asym=3, xmid = 0, scal = 1),
                 algorithm = "plinear", trace = TRUE))
if (! inherits(fm2a, "try-error")) summary(fm2a)
```
